---
title: "Central Limit Theorem"
author: "Mike Martos"
date: "May 28, 2016"
output: 
  html_document:
    keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Example 01: Starting from a normal distribution with a sample size 100
```{r NormalDistribution10}
#Original Sample
origSample <- rnorm(100, 0, 1)

#Show summary of original sampple
summary(origSample)

#Store measurements from original sample
origMean <- mean(origSample)
origMedian <- median(origSample)
origSD <- sd(origSample)
origVar <- var(origSample)

#Plot a histogram of Original Sample
hist(origSample)

#Start Iterarion process
nsim <- 1000
meanArray <- numeric(nsim)
medianArray <- numeric(nsim)

for (i in 1:nsim) {
  bootsample <- sample(origSample, size = 100, replace = TRUE)
  meanArray[i] <- mean(bootsample)
  medianArray[i] <- median(bootsample)
}

#Plot histogram of array of means after process
hist(meanArray)

#Plot lines with significant quantities Original Mean, New Mean, Quantiles
abline(v=origMean,col="red", lwd=2)
abline(v=mean(meanArray),col="blue", lty=3, lwd=2)
meanQuant <- quantile(meanArray, c(0.025, 0.975))
abline(v=meanQuant[1],col="green")
abline(v=meanQuant[2],col="green")

#Plot histogram of array of medians after process
hist(medianArray)

#Plot lines with significant quantities Original Mean, New Mean, Quantiles
abline(v=origMedian,col="red", lwd=2)
abline(v=median(medianArray),col="blue", lty=3, lwd=2)
meanQuant <- quantile(medianArray, c(0.025, 0.975))
abline(v=meanQuant[1],col="green")
abline(v=meanQuant[2],col="green")

#Show Original statistics
origMean
origMedian
origSD
origVar

#Show New statistics
mean(meanArray)
median(medianArray)
sd(medianArray)
var(medianArray)

# Show boxplots side by side
lmts <- range(origSample, meanArray)
par(mfrow = c(1, 2))
boxplot(origSample, ylim=lmts, xlab="Original Sample")
boxplot(meanArray, ylim=lmts, xlab="After bootsrap process")
```

# Example 02: Starting from a normal distribution with a sample size 10
```{r NormalDistribution50}
#Original Sample
origSample <- rnorm(10, 0, 1)

#Show summary of original sampple
summary(origSample)

#Store measurements from original sample
origMean <- mean(origSample)
origMedian <- median(origSample)
origSD <- sd(origSample)
origVar <- var(origSample)

#Plot a histogram of Original Sample
hist(origSample)

#Start Iterarion process
nsim <- 1000
meanArray <- numeric(nsim)
medianArray <- numeric(nsim)

for (i in 1:nsim) {
  bootsample <- sample(origSample, size = 10, replace = TRUE)
  meanArray[i] <- mean(bootsample)
  medianArray[i] <- median(bootsample)
}

#Plot histogram of array of means after process
hist(meanArray)

#Plot lines with significant quantities Original Mean, New Mean, Quantiles
abline(v=origMean,col="red", lwd=2)
abline(v=mean(meanArray),col="blue", lty=3, lwd=2)
meanQuant <- quantile(meanArray, c(0.025, 0.975))
abline(v=meanQuant[1],col="green")
abline(v=meanQuant[2],col="green")

#Plot histogram of array of medians after process
hist(medianArray)

#Plot lines with significant quantities Original Mean, New Mean, Quantiles
abline(v=origMedian,col="red", lwd=2)
abline(v=median(medianArray),col="blue", lty=3, lwd=2)
meanQuant <- quantile(medianArray, c(0.025, 0.975))
abline(v=meanQuant[1],col="green")
abline(v=meanQuant[2],col="green")

#Show Original statistics
origMean
origMedian
origSD
origVar

#Show New statistics
mean(meanArray)
median(medianArray)
sd(medianArray)
var(medianArray)

# Show boxplots side by side
lmts <- range(origSample, meanArray)
par(mfrow = c(1, 2))
boxplot(origSample, ylim=lmts, xlab="Original Sample")
boxplot(meanArray, ylim=lmts, xlab="After bootsrap process")
```

# Example 03: Starting from a normal distribution with a sample size 50
```{r ExpDistribution50}
#Original Sample
origSample <- rexp(50)

#Show summary of original sampple
summary(origSample)

#Store measurements from original sample
origMean <- mean(origSample)
origMedian <- median(origSample)
origSD <- sd(origSample)
origVar <- var(origSample)

#Plot a histogram of Original Sample
hist(origSample)

#Start Iterarion process
nsim <- 1000
meanArray <- numeric(nsim)
medianArray <- numeric(nsim)

for (i in 1:nsim) {
  bootsample <- sample(origSample, size = 50, replace = TRUE)
  meanArray[i] <- mean(bootsample)
  medianArray[i] <- median(bootsample)
}

#Plot histogram of array of means after process
hist(meanArray)

#Plot lines with significant quantities Original Mean, New Mean, Quantiles
abline(v=origMean,col="red", lwd=2)
abline(v=mean(meanArray),col="blue", lty=3, lwd=2)
meanQuant <- quantile(meanArray, c(0.025, 0.975))
abline(v=meanQuant[1],col="green")
abline(v=meanQuant[2],col="green")


#Plot histogram of array of medians after process
hist(medianArray)

#Plot lines with significant quantities Original Mean, New Mean, Quantiles
abline(v=origMedian,col="red", lwd=2)
abline(v=median(medianArray),col="blue", lty=3, lwd=2)
meanQuant <- quantile(medianArray, c(0.025, 0.975))
abline(v=meanQuant[1],col="green")
abline(v=meanQuant[2],col="green")

#Show Original statistics
origMean
origMedian
origSD
origVar

#Show New statisticsmean(meanArray)
median(medianArray)
sd(medianArray)
var(medianArray)

# Show boxplots side by side
lmts <- range(origSample, meanArray)
par(mfrow = c(1, 2))
boxplot(origSample, ylim=lmts, xlab="Original Sample")
boxplot(meanArray, ylim=lmts, xlab="After bootsrap process")

```

# Example 04: Starting from a normal distribution with a sample size 10 
```{r ExpDistribution10}
#Original Sample
origSample <- rexp(10)

#Show summary of original sampple
summary(origSample)

#Store measurements from original sample
origMean <- mean(origSample)
origMedian <- median(origSample)
origSD <- sd(origSample)
origVar <- var(origSample)

#Plot a histogram of Original Sample
hist(origSample)

#Start Iterarion process
nsim <- 1000
meanArray <- numeric(nsim)
medianArray <- numeric(nsim)

for (i in 1:nsim) {
  bootsample <- sample(origSample, size = 10, replace = TRUE)
  meanArray[i] <- mean(bootsample)
  medianArray[i] <- median(bootsample)
}

#Plot histogram of array of means after process
hist(meanArray)

#Plot lines with significant quantities Original Mean, New Mean, Quantiles
abline(v=origMean,col="red", lwd=2)
abline(v=mean(meanArray),col="blue", lty=3, lwd=2)
meanQuant <- quantile(meanArray, c(0.025, 0.975))
abline(v=meanQuant[1],col="green")
abline(v=meanQuant[2],col="green")

#Plot histogram of array of means after process
hist(medianArray)

#Plot lines with significant quantities Original Mean, New Mean, Quantiles
abline(v=origMedian,col="red", lwd=2)
abline(v=median(medianArray),col="blue", lty=3, lwd=2)
meanQuant <- quantile(medianArray, c(0.025, 0.975))
abline(v=meanQuant[1],col="green")
abline(v=meanQuant[2],col="green")

#Show Original statistics
origMean
origMedian
origSD
origVar

#Show New statisticsmean(meanArray)
median(medianArray)
sd(medianArray)
var(medianArray)


# Show boxplots side by side
lmts <- range(origSample, meanArray)
par(mfrow = c(1, 2))
boxplot(origSample, ylim=lmts, xlab="Original Sample")
boxplot(meanArray, ylim=lmts, xlab="After bootsrap process")

```

#Observations:

I've used two measures of center (*Mean and Median*) to graph the differences, knowing that the Central limit theorem involves the *mean* and not the *median*, and also that the bootstrapping process works well with statistics that involve the whole data set.

My intention was to observe graphically, the behavior of both measures of center, a resistant and a nonresistant one.

Examples 01 and 02 start from an already normal distributed sample and the last two from an exponential sample.


###After visual observation of the output of this code, I could identify several things:

* With more observations in a sample the likelihood of getting **larger observations** is higher, on both, normal distribution and exponential distribution.

* On examples starting from a normal distribution already, the spread gets tighter after the bootstrap process, it is still nearly normal but the variance and standard error become smaller.

* On examples 03 and 04 that start with an exponential distribution, after the bootstrap process the means of the samples show a nearly normal distribution.

* *Means* before and after the bootstrap process are very close to each other.

* *Medians* before and after seem to be the same every time I've ran this code.

* The bootstrapping process works well with *means* but not so well with *medians*, the distributions are not normal, which is expected as the ***Central Limit Theorem*** is based on the *mean*.









